---
title: "The Sprawl Tax: Current Atlanta Zoning Map"
output: html_document
date: "2025-03-26"
---
## 🗺️ How to Use This Map

- Use the **layer control** (top-right corner of the map) to toggle:
  - **Zoning categories**: Single Family, Multi Family, Commercial.
  - **Economic overlays**: Median Income, Median Home Value.
  - **Social overlays**: % Black, % White, % Asian, % Hispanic.
- You can turn **multiple layers on or off** at the same time.
- Click on areas for more detailed information.
- Zoom and pan the map to explore specific neighborhoods in Atlanta.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE)
```
```{r}
library(ggplot2)
library(sf)
library(dplyr)
library(leaflet)
library(leaflet.extras)
library(tidycensus)
library(tidyr)
library(tigris)
library(leaflet.extras2)
census_api_key(Sys.getenv("CENSUS_API_KEY"), install = FALSE)
options(tigris_use_cache = TRUE)

zoning_atl_json <- st_read("/Users/ryannation/Downloads/Official_Lots_with_Zoning_.geojson", quiet = TRUE)
zoning_atl_csv <- read.csv("/Users/ryannation/Downloads/Official_Lots_with_Zoning_.csv")

ga_places <- places(state = "GA", year = 2021)
atlanta_city_limit <- ga_places %>% filter(NAME == "Atlanta")

acs_dem <- c(
  median_income = "B19013_001")
atlanta_acs <- get_acs(
  geography = "tract",
  variables = acs_dem,
  state = "GA",
  county = "Fulton",
  geometry = TRUE,
  year = 2021)

merged_atl <- zoning_atl_json %>%
  left_join(zoning_atl_csv, by = "OBJECTID_12") %>%
  rename(ZONINGCODE = ZONINGCODE.x) %>%
  rename(AREA = AREA.x) %>%
  select(OBJECTID_12, ZONINGCODE, AREA, geometry)

merged_atl <-st_zm(merged_atl, drop = TRUE, what = "ZM")

single_family <- "Single Family Residential"
multi_family <- "Multi-Family Residential"
commercial_types <- c("Commercial", "Industrial", "Special Public Interest")

merged_atl <- merged_atl %>%
  mutate(ZONING_CATEGORY = case_when(
    ZONINGCODE == single_family ~ "Single Family Residential",
    ZONINGCODE == multi_family ~ "Multi-Family Residential",
    ZONINGCODE %in% commercial_types ~ "Commercial",
    TRUE ~ NA_character_)) %>%
  filter(!is.na(ZONING_CATEGORY))

filtered_zones <- merged_atl %>%
  filter(st_is_valid(geometry)) %>%
  group_by(ZONING_CATEGORY) %>%
  summarize(geometry = st_union(geometry), .groups = "drop")

zoning_pal <- colorFactor(
  palette = c("black", "black", "black"),
  domain = filtered_zones$ZONING_CATEGORY)

atlanta_acs <- st_transform(atlanta_acs, crs = st_crs(filtered_zones))
atlanta_city_limit <- st_transform(atlanta_city_limit, crs = st_crs(atlanta_acs))

income_data <- atlanta_acs %>%
  filter(variable == "median_income") %>%
  filter(st_intersects(., atlanta_city_limit, sparse = FALSE))
income_pal <- colorNumeric(palette = "Oranges", domain = income_data$estimate)

race_pct <- get_acs(
  geography = "tract",
  variables = c(
    total = "B02001_001",
    black = "B02001_003",
    white = "B02001_002",
    asian = "B02001_005",
    hispanic = "B03003_003"),
  state = "GA",
  county = "Fulton",
  geometry = TRUE,
  year = 2021) %>%
  select(GEOID, variable, estimate, geometry) %>%
  st_transform(st_crs(filtered_zones))

race_pct_wide <- race_pct %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = variable, values_from = estimate) %>%
  mutate(
    pct_black = (black / total) * 100,
    pct_white = (white / total) * 100,
    pct_asian = (asian / total) * 100,
    pct_hispanic = (hispanic / total) * 100)
geometry_data <- race_pct %>% distinct(GEOID, geometry)
race_pct_wide <- left_join(race_pct_wide, geometry_data, by = "GEOID")
race_pct_wide <- st_as_sf(race_pct_wide) %>%
  filter(st_intersects(., atlanta_city_limit, sparse = FALSE))

black_palette <- colorNumeric("Greens", domain = race_pct_wide$pct_black)
white_palette <- colorNumeric("Blues", domain = race_pct_wide$pct_white)
asian_palette <- colorNumeric("Purples", domain = race_pct_wide$pct_asian)
hispanic_palette <- colorNumeric("BuPu", domain = race_pct_wide$pct_hispanic)

home_value_var <- c(median_home_value = "B25077_001")
home_value_data <- get_acs(
  geography = "tract",
  variables = home_value_var,
  state = "GA",
  county = "Fulton",
  geometry = TRUE,
  year = 2021)

home_palette <- colorNumeric("Reds", domain = home_value_data$estimate, na.color = "transparent")

race_data <- st_transform(race_pct_wide, crs = st_crs(filtered_zones))
home_value_data <- st_transform(home_value_data, crs = st_crs(atlanta_acs))
home_value_data <- home_value_data %>%
  filter(st_intersects(., atlanta_city_limit, sparse = FALSE)) %>%
  st_transform(home_value_data, crs = st_crs(filtered_zones))

atl_map <- leaflet(filtered_zones) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolylines(
    data = st_transform(atlanta_city_limit, crs = st_crs(filtered_zones)),
    color = "black",
    weight = 3,
    label = "City of Atlanta",
    group = "City Limits") %>%
  addPolygons(
    data = filtered_zones %>% filter(ZONING_CATEGORY == "Single Family Residential"),
    fillColor = ~zoning_pal(ZONING_CATEGORY),
    fillOpacity = 1,
    color = NA,
    weight = 0,
    label = ~paste("Zoning:", ZONING_CATEGORY),
    group = "Zoning: Single Family") %>%
  addPolygons(
    data = filtered_zones %>% filter(ZONING_CATEGORY == "Multi-Family Residential"),
    fillColor = ~zoning_pal(ZONING_CATEGORY),
    fillOpacity = 1,
    color = NA,
    weight = 0,
    label = ~paste("Zoning:", ZONING_CATEGORY),
    group = "Zoning: Multi Family") %>%
  addPolygons(
    data = filtered_zones %>% filter(ZONING_CATEGORY == "Commercial"),
    fillColor = ~zoning_pal(ZONING_CATEGORY),
    fillOpacity = 1,
    color = NA,
    weight = 0,
    label = ~paste("Zoning:", ZONING_CATEGORY),
    group = "Zoning: Commercial") %>%
  addPolygons(
    data = home_value_data,
    fillColor = ~home_palette(estimate),
    fillOpacity = 0.6,
    color = "black",
    weight = 0.4,
    label = ~paste0("Median Home Value: $", formatC(estimate, format = "f", digits = 0, big.mark = ",")),
    group = "Economic: Median Home Value") %>%
  addPolygons(
    data = income_data,
    fillColor = ~income_pal(estimate),
    fillOpacity = 0.6,
    color = "black",
    weight = 0.4,
    label = ~paste0("Median Income: $", formatC(estimate, format = "f", digits = 0, big.mark = ",")),
    group = "Economic: Median Income") %>%
  addPolygons(
    data = race_pct_wide,
    fillColor = ~black_palette(pct_black),
    fillOpacity = 0.6,
    color = "black",
    weight = 0.4,
    label = ~paste0("Black: ", round(pct_black, 1), "%"),
    group = "Social: % Black") %>%
  addPolygons(
    data = race_pct_wide,
    fillColor = ~white_palette(pct_white),
    fillOpacity = 0.6,
    color = "black",
    weight = 0.4,
    label = ~paste0("White: ", round(pct_white, 1), "%"),
    group = "Social: % White") %>%
  addPolygons(
    data = race_pct_wide,
    fillColor = ~asian_palette(pct_asian),
    fillOpacity = 0.6,
    color = "black",
    weight = 0.4,
    label = ~paste0("Asian: ", round(pct_asian, 1), "%"),
    group = "Social: % Asian") %>%
  addPolygons(
    data = race_pct_wide,
    fillColor = ~hispanic_palette(pct_hispanic),
    fillOpacity = 0.6,
    color = "black",
    weight = 0.4,
    label = ~paste0("Hispanic: ", round(pct_hispanic, 1), "%"),
    group = "Social: % Hispanic") %>%
  addLayersControl(
    baseGroups = c("Zoning: Single Family", "Zoning: Multi Family", "Zoning: Commercial"),
    overlayGroups = c("Economic: Median Home Value", "Economic: Median Income", 
      "Social: % Black", "Social: % White", "Social: % Asian", "Social: % Hispanic"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("Economic: Median Home Value", "Economic: Median Income", 
      "Social: % Black", "Social: % White", "Social: % Asian", "Social: % Hispanic")) %>%
  setView(lng = -84.388, lat = 33.748, zoom = 10.7)

atl_map
```
